### 2024-11-04 paper.tex
add paper

\documentclass{article}
\usepackage{graphicx} % Required for inserting images

\title{Analyzing citation rate of a scientific paper using ML and graph algorithms}
\author{Vladimir Zakharov, Sergey Fedchin, Artem Chebykin}
\date{Fall 2024}

\begin{document}

\maketitle

\section{Introduction}

\end{document}